Open-source landscape for 3D CSEM modelling

### 2020-09-11 manuscript/Manuscript.tex
Updates from Overleaf
Move local up to 2nd version

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most common applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or OcTree meshes \citep{ECP.07.Haber}.

@@

Meshes and finite volume differential operators are implemented in the \discretize package, which currently includes tensor, OcTree, cylindrical, and logically rectangular meshes. Each mesh type inherits from a common structure and uses the same naming conventions for methods and properties. This allows us to decouple the implementation of a discretized partial differential equation from the details of the mesh geometry and therefore we can write a single implementation of discretized partial differential equation (or set of partial differential equations) in \simpeg that will support all mesh types implemented in \discretize.

@@

\itodo{HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE}
We computed the responses for two different models to verify that all different 3D codes yield the same electromagnetic responses. The first simple \emph{Block model} consists of a layered, anisotropic background with three embedded blocks. The second model, \emph{Marlim R3D}, is based on a very complex, realistic marine CSEM setup.

@@

\imycom[DW]{There are some interesting features we should highlight. But again,
we first need all results. E.g.}
\begin{itemize}
  \item 0.125\,Hz: on the left side, \custem and \emg3d are almost identical,
    but very different from MR3D. So there I actually trust more our
    calculation, or there is a difference in the model that we did not take
    into account.
  \item 0.125\,Hz: on the right side, \custem is very similar to MR3D, whereas
    \emg3d is off.
  \item 0.125\,Hz and 0.5\,Hz: on the right side, \custem seems to very similar
    to MR3D.
\end{itemize}

%
\tabl[btp]{comp-marlim}{Comparison of used runtime, memory, and number of cells
  used to discretize the Marlim R3D model.}{
  \centering
\begin{tabular}{lS[table-format=6.0]rS[table-format=4.1]S[table-format=8.0]}
  \toprule
  %
       & {Runtime (s)} & \#Procs & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 4984 & 64 & 558.5 & 4539474 \\ % dof: 9078948 / 2 (Real+Imag -> Complex)
  \emg3d  & 1478 &  1 &   0.5 & 5998992 \\
  \petgem & 3527 & 96 & 437.8 & 4539474 \\
  \simpeg &  422 &  4 &  12.8 &  720146 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

@@

The computation times can differ significantly for specific simulations, but there is no clear general advantage for one of the solver types. In our models, the runtime comparison favoured the iterative solver and therefore, \emg3d. Even though not shown here, direct solvers exhibit their strongest advantage in terms of computation times if responses of multiple CSEM transmitters need to be calculated in the same computational domain. As the system matrix factorisation requires 98-99\,\% of the solution time, computations for additional sources come at almost no cost for direct solvers, i.e., MUMPS, whereas it would come at the same cost as the first source for \emg3d. For the sake of completeness, note that there are also advanced techniques for iterative solvers to make use of preconditioners, factorisations, or intermediate solutions of the previous SLE to speed up the calculation for multiple sources, but this argumentation is beyond the scope of this work.

@@

Within SimPEG: This work has motivated feature development including: (a) interpolation and averaging strategies from mapping physical properties on a fine mesh to a coarser mesh for computation and (b) new examples to include in the documentation for designing OcTree meshes. Furthermore, as a part of a broader development objective of inter-operating with other forward-simulation engines, connecting \emg3d and \simpeg provided a motivating use-case for the latest refactor and release of \simpeg 0.14.0.
% \citep{simpeg0.14.0}.
}

>>>>

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most common applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or octree meshes \citep{ECP.07.Haber}.

@@

Meshes and finite volume differential operators are implemented in the \discretize package, which currently includes tensor, octree, cylindrical, and logically rectangular meshes. Each mesh type inherits from a common structure and uses the same naming conventions for methods and properties. This allows us to decouple the implementation of a discretized partial differential equation from the details of the mesh geometry and therefore we can write a single implementation of discretized partial differential equation (or set of partial differential equations) in \simpeg that will support all mesh types implemented in \discretize.

@@

We computed the responses for three different models to verify that the four 3D CSEM codes yield the same electromagnetic responses and to compare different solver types (FE, FD) and different mesh types (unstructured tetrahedra, tensor, octree) in different scenarios. The first model is an anisotropic layered model, which can be compared to semi-analytical solutions. The layered model serves also as background model for the second verification, where we add three resistive blocks into the subsurface. The third test is based on a realistic, anisotropic marine model.

\subsection{Layered Model}

\itodo{HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE HERE}

@@

\imycom[DW]{There are some interesting features we should highlight. But again,
we first need all results. E.g.}
\begin{itemize}
  \item 0.125\,Hz: on the left side, \custem and \emg3d are almost identical,
    but very different from MR3D. So there I actually trust more our
    computation, or there is a difference in the model that we did not take
    into account.
  \item 0.125\,Hz: on the right side, \custem is very similar to MR3D, whereas
    \emg3d is off.
  \item 0.125\,Hz and 0.5\,Hz: on the right side, \custem seems to very similar
    to MR3D.
  \item FD needs comparably a lot of cells
\end{itemize}

%
\tabl[btp]{comp-marlim}{Comparison of used runtime, memory, and number of cells
  used to discretize the Marlim R3D model.}{
  \centering
\begin{tabular}{lS[table-format=6.0]rS[table-format=4.1]S[table-format=8.0]}
  \toprule
  %
       & {Runtime (s)} & \#Procs & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 4984 & 64 & 558.5 & 4539474 \\ % dof: 9078948 / 2 (Real+Imag -> Complex)
  \emg3d  & 1200 &  1 &   0.5 & 5998992 \\
  \petgem & 3527 & 96 & 437.8 & 4539474 \\
  \simpeg &  422 &  4 &  12.8 &  720146 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

@@

The computation times can differ significantly for specific simulations, but there is no clear general advantage for one of the solver types. In our models, the runtime comparison favoured the iterative solver and therefore, \emg3d. Even though not shown here, direct solvers exhibit their strongest advantage in terms of computation times if responses of multiple CSEM transmitters need to be computed in the same computational domain. As the system matrix factorisation requires 98-99\,\% of the solution time, computations for additional sources come at almost no cost for direct solvers, i.e., MUMPS, whereas it would come at the same cost as the first source for \emg3d. For the sake of completeness, note that there are also advanced techniques for iterative solvers to make use of preconditioners, factorisations, or intermediate solutions of the previous SLE to speed up the computation for multiple sources, but this argumentation is beyond the scope of this work.

@@

Within SimPEG: This work has motivated feature development including: (a) interpolation and averaging strategies from mapping physical properties on a fine mesh to a coarser mesh for computation and (b) new examples to include in the documentation for designing octree meshes. Furthermore, as a part of a broader development objective of inter-operating with other forward-simulation engines, connecting \emg3d and \simpeg provided a motivating use-case for the latest refactor and release of \simpeg 0.14.0.
% \citep{simpeg0.14.0}.
}
